How Short Lived is the Population III Epoch?

### 2010-09-01 ms.tex
Adding to the star formation section.

\shorttitle{TRANSITION TO POPULATION II}
\shortauthors{WISE ET AL.}

@@

We use the adaptive mesh refinement (AMR) code \enzo~\citep{refs} that
employs our implementation of adaptive ray tracing for radiation
transport \citep{refs}.  To resolve minihalos with at least 100 dark
matter (DM) particles and follow the formation of the first generation
of dwarf galaxies, we use a simulation box of 1 Mpc that has a
resolution of $256^3$.  This gives us a DM mass resolution of 1840
\Ms.  We refine the grid on baryon overdensities of $3 \times
2^{-0.2l}$, where $l$ is the AMR level, resulting in a
super-Lagrangian behavior.  We also refine on a DM overdensity of $3
\times 2^l$ and always resolve the local Jeans length by at least 4
cells.  The latter criterion avoids any artifical fragmentation during
gaseous collapses \citep{Truelove97}.  We initialize the simulation
with \textsl{grafic} \citep{Bertschinger01} at $z = ???$ and use the
cosmological parameters from the 7-year WMAP data \citep{ref}:
$\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h
= 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$ with the variables having
their usual definitions.

We peform two simulations with the same initial conditions.  They
differ only in that one considers primordial cooling, and the other
considers additional cooling in metal-enriched gas.  For primordial
cooling, we use a non-equilbrium chemistry solver with 9 species of
hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
cooling curve that is calculated from CLOUDY \citep{ref} and using the
method of \citet{ref}.  It is tabulated based on density, temperature,
electron fraction, and metallicity.

@@

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{Turk10} has recently
shown that Pop III binaries may form in a fraction of such halos.
Furthermore, detailed one-dimensional calculations have shown that
their IMF should follow a Kroupa-like IMF that has a Salpeter slope at
the high-mass end and an exponential cut-off below $M_{\rm char}$
\citep{refs}.

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} but now the stellar mass is
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = ...
\end{equation}
%
For reproducibility, we initialize both simulations with the same
random seed and record the number of times the random number generator
\citep[Mersenne twister;][]{ref} has been called.

>>>>

\shorttitle{POPULATION III STAR EPOCH}
\shortauthors{WISE ET AL.}

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise11}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = ???$ and use the cosmological parameters from the 7-year WMAP
data \citep{ref}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

We peform two simulations with the same initial conditions.  They
differ only in that one considers primordial cooling, and the other
considers additional cooling in metal-enriched gas.  For primordial
cooling, we use a non-equilbrium chemistry solver with 9 species of
hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
cooling curve that is calculated from CLOUDY \citep{ref} and using the
method of \citet{2008MNRAS.385.1443S}.  It is tabulated based on
density, temperature, electron fraction, and metallicity.

@@

We distinguish Pop II and Pop III star formation by the metallicity of
the cold ($T < 10^3$ K) star forming gas in the molecular clouds.  Pop
II stars are formed if [Z/H] $> -4$, and Pop III stars are formed
otherwise.  We do not consider Pop III.2 stars and intermediate mass
stars from CMB-limited cooling.

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{2009Sci...325..601T}
has recently shown that Pop III binaries may form in a fraction of
such halos.  Furthermore, detailed one-dimensional calculations have
shown that their IMF should follow a Kroupa-like IMF that has a
Salpeter slope at the high-mass end and an exponential cut-off below
$M_{\rm char}$ \citep{refs}.

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
%
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{ref} has been
called for use when restarting the simulations.
